Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-08-01 demystify.tex
updates

\subsection{Selecting and Ranking Schools}\label{sec:selecting-ranking-schools}
\begin{center}
  \includegraphics[scale=0.5]{files/c1.png}
\end{center}

International students often put schools into two bins: (i) very top schools that they dream about, and (ii) everything else.  They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.  Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
They even rank universities based on popular states they know in the US, e.g., California, Texas, and New York.

You can learn about CS programs and the research interests of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}. You will be very surprised to learn that a school that you didn't know much about can have very strong research in your topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in). This is also a good way to learn about individual faculty, e.g., who works on what, and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research findings (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}. %\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.

@@

\subsubsection*{Additional Resources}
\begin{itemize}
  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive ``what do I ask/look for'' in a PhD Advisor Guide}
\end{itemize}
\subsection{Types of Faculty: Who can serve as a PhD adviser?}\label{sec:faculty-types}

@@

Tenure-track faculty such as assistant professors are likely young and very active in research (they have to, to get tenure). Thus, they will likely have more time to work with you and push you to do research and publish. However, they may not have as much experience in managing students and may not have as much funding (yet).

@@

@@

Also thanks to NSF its flexibility and allows faculty to be creative in research and education, which allows me to have time to write this document.

>>>>

\subsection{Selecting and Ranking Schools}\label{sec:selecting-ranking-schools}
\begin{center}
  \includegraphics[scale=0.6]{files/c1.png}
\end{center}

International students often put schools into two bins: (i) top schools that they dream about, and (ii) everything else.  They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.  Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
They even rank universities based on popular places they know in the US, e.g., California, Texas, and New York.

Instead of these superficial criteria, you should consider the CS program and the research interests of faculty members.
You can learn about these using resources such as \href{https://csrankings.org}{CSRankings.org}. You will be very surprised to learn that a school that you didn't know much about can have very strong research in your topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in). This is also a good way to learn about individual faculty, e.g., who works on what, and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research findings (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}. %\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.

@@

\subsubsection*{Additional Resources}
\begin{itemize}
  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive ``what do I ask/look for'' in a PhD Advisor Guide}
\end{itemize}
\subsection{Who can serve as a PhD adviser?}\label{sec:faculty-types}

@@

Now that you know a bit about tenured and tenure-track faculty (\autoref{sec:faculty-types}).  Which one should you choose as your adviser? The short answer is that either can be a good adviser, but they have different strengths and weaknesses.

Tenure-track faculty, e.g., assistant professors, are likely young and active in research (they have to, to get tenure). Thus, they will devote more time to work with you and push you to do research and publish. However, they may not have as much experience in managing students and may not have as much funding (yet).

@@

\subsection{Things to avoid when cold emailing a professor}

\begin{itemize}

    \item \textbf{Not customizing the email}. Many students write a generic email and send it to multiple professors. Tthese emails typically talk about the student and not why they are interested in the professor's work. See writing tips in \autoref{sec:contact}
    \item \textbf{Sending the same email to multiple professors}. Some students send to multiple emails using BCC or even CC (where we can see all addresses the email being sent to). This is a big turn-off for professors and will result in no reply or a very harsh reply on how unprofessional you are.
    \item \textbf{Too much information}. Keep your email short and to the point. If we are interested, we will ask for more information.
    \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email.  Use Prof. Lastname or Dr. Lastname (\autoref{sec:address}).
    \item \textbf{Not reading the professor's website} and asking questions that are already answered there.
\end{itemize}

@@

Also thanks to NSF for encouraging faculty to be creative in research and education, which allows me to write this document.
